Simulating Generalized Fading Channels in Python

### 2018-05-14 paper/manuscript.tex
update paper

\documentclass[conference, 10pt]{IEEEtran}
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{color}
\usepackage{placeins}
\usepackage{bm}
\usepackage{cite}
%\usepackage{stfloats}
\usepackage{float}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{tabularx,colortbl}
\DeclareMathOperator*{\argmin}{arg\,min}

\begin{document}
\title{\texttt{maoud}: a Python package for Simulating Generalized Fading Channels}

\author{\IEEEauthorblockN{Jos\'e~V.~de~M.~Cardoso,~Paulo~R.~Lins~J\'unior,}
        \IEEEauthorblockN{~Wamberto~J.~L.~Queiroz,~and~Marcelo~S.~Alencar,~\textit{IEEE Senior Member}}
\IEEEauthorblockA{Universidade Federal de Campina Grande\\
Instituto Federal de Educa\c c\~ao, Ciencia, e Tecnologia da Para\'iba\\
Campina Grande, Para\'iba, Brasil\\
\{josevinicius, wamberto, paulo, malencar\}@iecom.org.br\\}
}

@@

\begin{abstract}
    We present a well tested Python-based library for simulating and computing
    generalized fading channels, named \texttt{maoud}. We describe the
    applicability of \texttt{maoud} using examples in scenarios of communications
    channels impaired by generalized fading, namely: spectrum sensing and bit error
    rate computation. The development of \texttt{maoud} is open source and its code,
    along with the code for the examples presented in this paper, are avaliable at
    \texttt{http://github.com/mirca/maoud}.
\end{abstract}

\IEEEpeerreviewmaketitle
\section{Introduction}

@@

Although accurate and precise distributions for generalized fading have been
estabilished in the literature, such as $\alpha\text{-}\mu$~\cite{yacoub.2007.1},
$\kappa\text{-}\mu$~\cite{yacoub.2007.2} and $\eta\text{-}\mu$~\cite{yacoub.2007.2},
the generation of samples following these distributions is usually
a time-consuming task. In~\cite{cogliatti2012}, the authors built an efficient
algorithm, based on the rejection method, for generation of samples from those
distributions.

however, there are neither open nor
closed source implementations available to the scientific community.

In this paper, we present an open source Python package, named \texttt{maoud},
for generation of samples following the $\alpha\text{-}\mu$, $\kappa\text{-}\mu$, and
$\eta\text{-}\mu$ distributions. The usefullness of $\texttt{maoud}$ is illustrated
through examples involving spectrum sensing, bit error rate (BER) computation, and
fading estimation.

\subsection*{Notation}

\section{Rejection sampling}

\section{Examples}
\subsection{Spectrum Sensing in Complex Generalized Fading Channels}

@@

From a probabilistic point of view, the spectrum sensing problem may be framed as
a decision theory problem, as follows
\begin{align}
    H_0:~& \bm{y} = \bm{w},\\
    H_1:~& \bm{y} = h\bm{s} + \bm{w},
\end{align}
in which $\bm{y} \in \mathbb{C}^{n\times 1}$ is the received vector signal,
$\bm{w} \in \mathbb{C}^{n\times 1}$ is complex Gaussian noise process with zero mean
vector and covariance matrix given as $\sigma^2\bm{I}_n$, and $h$ is the channel gain.

@@

Recall that the energy detection rule can be expressed as
\begin{equation}
    d_\delta (\tilde{y}) = \mathbb{I}(\tilde{y} > \delta)
\end{equation}
in which $\delta$ is a strictly positive real number known as energy threshold,
and $d_\delta (\tilde{y}) = j,~j \in \{0,1\}$, means that the detector has decided
in favor of the hypothesis $H_j$.

@@

The performance of detection schemes can be measured by computing the Receiver Operating
Characteristic (ROC), which consists in varying $\delta$ and computing the pairs of
probability of false alarm and miss detection, as illustrated in Fig.~\ref{fig:spectrum-sensing}.

\begin{figure}[!htb]
    \centering
    \includegraphics{figures/spectrum_sensing.eps}
    \caption{Receiver Operating Characteristic for the energy detector in $\alpha$-$\mu$
             fading channel with $\alpha=2$ and $\mu=1$, i.e., Nakagami-$m$, $m = 1$.
             The solid curve represents the theorectical probabilities as stated on
             (\ref{eq:pf}) and (\ref{eq:pd}) for different values of
             $\delta$, whereas bullets represent Monte Carlo simulations acquired with $10^6$ realizations.
             The input signal $\bm{s}$ consists in a vector of length $n=25$ in which each entry represents
             a symbol of a 64-PSK constellation. Symbols are assumed to be equiprobable.
             The signal-to-noise ratio is set to $5$ dB.}
    \label{fig:spectrum-sensing}
\end{figure}

\subsection{BER in $\alpha$-$\mu$ Fading}

@@

\begin{figure}[!htb]
    \centering
    \includegraphics{figures/ber.eps}
    \caption{BER}
    \label{fig:ber}
\end{figure}

\section{Conclusions}

\bibliographystyle{IEEEtran}
\bibliography{manuscript.bib}

>>>>

\documentclass{sbrt2013port}
%\usepackage[brazil]{babel}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{color}
%\usepackage{placeins}
\usepackage{bm}
%\usepackage{cite}
%\usepackage{stfloats}
%\usepackage{float}
\usepackage{hyperref}
%\usepackage{cite}
\usepackage{tabularx,colortbl}

\begin{document}

\title{Simulating Generalized Fading Channels in Python}

\author{José Vinícius de M. Cardoso$^1$,   Paulo Ribeiro Lins Júnior$^2$, Wamberto Queiroz$^1$ e Marcelo Sampaio de Alencar$^3$
\thanks{\noindent Instituto de Estudos Avançados em Comunicações -- Iecom; $^1$Universidade Federal de Campina Grande -- UFCG; $^2$Grupo de Pesquisa em Comunicações e Processamento de Informação -- GComPI/Instituto Federal de Educação, Ciência e Tecnologia da Paraíba -- IFPB, Campus Campina Grande -- PB, $^3$Universidade Federal da Bahia -- UFBA, emails: $^1$\{josevinicius, wamberto\}@iecom.org.br, $^2$paulo.lins@ifpb.edu.br e $^3$malencar@iecom.org.br.}}

@@

\markboth{XXXVI SIMPÓSIO BRASILEIRO DE TELECOMUNICAÇÕES E PROCESSAMENTO DE SINAIS - SBrT'2018, 16-19 DE SETEMBRO DE 2018, CAMPINA GRANDE, PB} {XXXVI SIMPÓSIO BRASILEIRO DE TELECOMUNICAÇÕES E PROCESSAMENTO DE SINAIS - SBrT'2018, 16-19 DE SETEMBRO DE 2018, CAMPINA GRANDE, PB}

\begin{abstract}
    We present a novel and well tested Python-based library for simulating and computing
    generalized fading channels, considering $\alpha\text{-}\mu$, $\kappa\text{-}\mu$, and
$\eta\text{-}\mu$ distributions. We describe the
    applicability of this library using a few communications channels scenarios
    impaired by generalized fading, namely, spectrum sensing and bit error
    rate computation. The development of the library is open source and its code along
    with examples are available at
    \texttt{http://github.com/mirca/maoud}.
\end{abstract}

\begin{keywords}
Generalized Fading, Simulation, Python
\end{keywords}

\section{Introduction}

@@

Although accurate and precise distributions for generalized fading have been
estabilished in the literature, such as $\alpha\text{-}\mu$~\cite{yacoub.2007.1},
$\kappa\text{-}\mu$~\cite{yacoub.2007.2} and $\eta\text{-}\mu$~\cite{yacoub.2007.2},
the generation of samples following these distributions is usually
a time-consuming task.

In~\cite{cogliatti2012}, the authors built an efficient
algorithm, based on the rejection method, for generation of samples from those
distributions. However, there are neither open nor
closed source implementations available to the scientific community.

In this paper, we present an open source Python package
for generation of samples following the $\alpha\text{-}\mu$, $\kappa\text{-}\mu$, and
$\eta\text{-}\mu$ distributions. The usefullness of ths package is illustrated
through well known examples involving spectrum sensing and bit error rate (BER) computation.

Section~\ref{rejec} presents the well known rejection method which is used as
the basis for the proposed implementation. Section~\ref{sensing} and~\ref{ber}
demonstrates the features of the proposed package in spectrum sensing over complex
generalized fading channels and bit error rate computation on $\alpha$-$\mu$ fading.
Section~\ref{conclusions} states the final remarks.

\section{Rejection Sampling}\label{rejec}

The algorithm implemented in this work, originally proposed in~\cite{cogliatti2012},
is based on the rejection method, which is one of the most used for the generation
of samples of a probability distribution.

We briefly describe the algorithm as follows. Consider that we would like to simulate a
random variable $x$ with a probability density $f$. The basic idea is to find
an alternative probability density $g$, from which there exists an efficient procedure
to generate samples, e.g., inverse methods. Additionally, it is required that $g$
majorizes $f$, i.e., we assume that there exists a real number $c$ such that $f(x)/g(x) < c$.

Basically, the steps of the rejection algorithm are\cite{ross}:
\begin{enumerate}
	\item generate a sample $y$ from the probability density $g$;
    \item generate a sample $u$ from a uniform density with support $(0, 1)$;
	\item check $u \leq \frac{f(y)}{cg(y)}$; if true then accept $y$; otherwise, return to step 1.
\end{enumerate}

The implementation of this algorithm along with the code to efficiently generate
samples from the $\alpha$-$\mu$, $\eta$-$\mu$, and $\kappa$-$\mu$ distributions
is available at \url{http://github.com/mirca/maoud}.

In order to validate the implementation, we apply the routines to
two use cases in wireless communications: spectrum sensing and bit error rate computation.

\section{Spectrum Sensing in Complex Generalized Fading Channels}~\label{sensing}

@@

From a probabilistic point of view, the spectrum sensing problem may be framed as
a decision theory problem, as follows
\begin{align}
    H_0:  \bm{y} &= \bm{w},\\
    H_1:  \bm{y} &= h\bm{s} + \bm{w},
\end{align}
\noindent in which $\bm{y} \in \mathbb{C}^{n\times 1}$ is the received vector signal,
$\bm{w} \in \mathbb{C}^{n\times 1}$ is complex Gaussian noise process with zero mean
vector and covariance matrix given as $\sigma^2\bm{I}_n$, and $h$ is the channel gain.

@@

Recall that the energy detection rule can be expressed as
\begin{equation}
    d_\delta (\tilde{y}) = \mathbb{I}(\tilde{y} > \delta)
\end{equation}
in which $\mathbb{I}$ is the indicator function, $\delta$ is a strictly positive real
number known as energy threshold, and $d_\delta (\tilde{y}) = j,~j \in \{0,1\}$,
means that the detector has decided in favor of the hypothesis $H_j$.

@@

The performance of detection schemes can be measured by computing the Receiver Operating
Characteristic (ROC), which consists in varying $\delta$ and computing the pairs of
probability of false alarm and miss detection, as illustrated in Fig.~\ref{fig:spectrum-sensing}, in which the solid curve represents the theorectical probabilities as stated on
(\ref{eq:pf}) and (\ref{eq:pd}) for different values of
$\delta$, whereas bullets represent Monte Carlo simulations acquired with $10^6$ realizations.
The input signal $\bm{s}$ consists in a vector of length $n=25$ in which each entry represents
a symbol of a 64-PSK constellation. Symbols are assumed to be equiprobable and
the signal-to-noise ratio is set to $5$ dB.

\begin{figure}[!htb]
    \centering
    \includegraphics{figures/spectrum_sensing}
    \caption{Receiver Operating Characteristic for the energy detector in $\alpha$-$\mu$
             fading channel with $\alpha=2$ and $\mu=1$.
             }
    \label{fig:spectrum-sensing}
\end{figure}

\section{Bit Error Rate in $\alpha$-$\mu$ Fading}~\label{ber}

@@

Fig.~\ref{fig:ber} shows the expression of the BER as a function of the signal-to-noise ratio.
As we can see, the simulation presents an excellent agreement against theorectical results.

\begin{figure}[!htb]
    \centering
    \includegraphics{figures/ber}
    \caption{BER \textit{versus} SNR.}
    \label{fig:ber}
\end{figure}

\section{Conclusions}\label{conclusions}

This paper presents a novel Python-based library for generalized fading simulation.
The implementation is tested and validated in two use cases, namely,
spectrum sensing and BER computation.

The computational results show that the implementation presents a
satisfactory agreement when compared with the theorectical formulation.

We expect that the proposed library, which is open source,
will be usefull to researchers, professors, and students of
the field of wireless communications.

\bibliographystyle{IEEEtran}
\bibliography{manuscript}
